\Title

### 2021-06-05 latex/paper/main.tex
First draft of proof of Thm is finished, need to edit now.

    Our assumption that \((\partial g/\partial r)(\theta,r_0,0) \equiv 0\)
    implies that \((\partial g/\partial \hat{r})(\theta,r_0,0) \equiv 0\).
    The solution of \ref{eqn:r1-hat-prime} can therefore be obtained by
    quadrature, giving
    \[
        \hat{r}_1(\theta,r_0) = \int\limits_0^\theta b(\sigma,r_0)d\sigma
        ,
    \]
    with \(b(\theta,r_0) := (\partial g/\partial I)(\theta,r_0,0)\). 
    Since and \(\dot{\theta} > 0\) and \(b(\theta,r_0)\) is \(2\pi\)-periodic in
    \(\theta\), there is a well-defined Poincar\'{e} map describing
    how \(\hat{r}\) changes each time \(\theta\) increases by \(2\pi\).
    This map is 
    \begin{align*}
        P : \, ]\underline{r},\overline{r}[ &\to \mathbb{R}_{\geq 0} \\
        r_0 &\mapsto \hat{r}(2\pi,r_0,I)
        ,
    \end{align*}
    which expands to \(P(r_0) = r_0 + IS(r_0) + R(2\pi,r_0,I)\).
    In other words, we can understand the evolution of \(\hat{r}\) by studying
    the dynamics of the discrete-time system
    \begin{equation}\label{eqn:r-discrete}
        r_{n+1} = r_n + IS(r_n) + R(2\pi,r_n,I)
        ,
    \end{equation}
    with initial condition \(r_0\).
    Here, \(r_n\) corresponds to the distance along the line segment
    \(L := ]\underline{r},\overline{r}[ \, \times \{0\}\)
    when an orbit of \((r,\theta)\) hits \(L\) for the the \(n^\text{th}\) time,
    assuming the acrobot was initialized at \((r,\theta) = (r_0,0)\).

@@

Next we compute
\(b(\theta,r_0) = (\partial g/\partial I)(\theta, r_0, 0) = K a(\theta,r_0)\),
where
\begin{align*}
    K &:= \frac{\bar{q}_a \sqrt{30m^2g l^3}}{15}
    , \\
    a(\theta,r_0) &:= \frac{
        r_0 |\sin(\theta)| \left(
        5 c_{r_0} \cos(r_0 c_\theta) - 8 \cos(r_0c_\theta)^2 + 3
    \right)
    }{
    \sin(r_0)\sqrt{\cos(r_0c_\theta) - c_{r_0}}
    }
    .
\end{align*}
Notice that \(K\) is a positive constant which depends only on
\(m\),\(g\),\(l\), and \(\bar{q}_a\), and \(a(\theta,r_0)\) is \(2\pi\)-periodic
in \(\theta\).
This means that there exists \(\epsilon > 0\) such that 
\(S(r_0) := \int_0^{2\pi} b(\sigma,r_0)d\sigma \geq \epsilon\)
if and only if \(Q(r_0) := \int_0^{2\pi} a(\sigma,r_0) d\sigma\) is strictly
positive for all \(r_0 \in \, ]0,\pi[\).
Since \(a(\theta,r_0)\) is adimensional, we can numerically compute \(Q(r_0)\)
once for all acrobots.
The plot of \(Q(r_0)\) when \(r_0 \in [10^{-10}, \pi - 10^{-3}]\) is shown in
Figure \ref{fig:acrobot-Q}.
We see that it is strictly positive and monotonically increasing, with an
asymptote at \(r_0 = \pi\).
Simulations with smaller \(r_0\) result in an infinite integral error due to a
division by zero, yet we believe the \(Q(r_0)\) is in fact positive for all 
\(r_0 \in \, ]0,\pi[\).
\begin{figure}
    \centering
    % TODO: Show this plot
    %\includegraphics[width=0.8\linewidth]{}
    \caption{The plot of \(Q(r_0)\).}
    \label{fig:acrobot-Q}
\end{figure}

While \(Q(r_0) \not \geq \epsilon\) on \(]0,\pi[\), it is certainly bounded
below on \(]\omega,\pi[\) for each fixed \(\omega > 0\).
This means all conditions of Lemma \ref{lemma:energy-regulation-D}
are satisfied on the set 
\[
    D_\omega = \mathcal{O}_1 \backslash 
    \{(q_u,p_u) \in \SxR \mid E(q_u,p_u) \leq E(\omega,0)\}
    .
\]
Therefore, for each \(\omega \in \, ]0,\pi[\) there exists some \(I_1^\star\)
such that the acrobot gains/loses energy everywhere on \(D_\omega\).
To prove energy gain on all of \(\mathcal{O}_1\), we will show the origin is a
repeller if \(I > 0\), and that it is an asymptotically stable equilibrium
if \(I < 0\).
Recall the Poincar\'{e} section used in the proof of Lemama
\ref{lemma:energy-regulation-D}.
\textbf{TODO: Show the origin is a repeller/stable eqm via the poincare map}.

>>>>

    Our assumption that \((\partial g/\partial r)(\theta,r_0,0) \equiv 0\)
    implies that \((\partial g/\partial \hat{r})(\theta,r_0,0) \equiv 0\).
    The solution of \ref{eqn:r1-hat-prime} can therefore be obtained by
    quadrature, giving
    \[
        \hat{r}_1(\theta,r_0) = \int\limits_0^\theta b(\sigma,r_0)d\sigma
        ,
    \]
    with \(b(\theta,r_0) := (\partial g/\partial I)(\theta,r_0,0)\). 
    Since and \(\dot{\theta} > 0\) and \(b(\theta,r_0)\) is \(2\pi\)-periodic in
    \(\theta\), there is a well-defined Poincar\'{e} map describing
    how \(\hat{r}\) changes each time \(\theta\) increases by \(2\pi\).
    This map is 
    \begin{align}\label{eqn:poincare-map}
        P : \, ]\underline{r},\overline{r}[ &\to \mathbb{R}_{\geq 0} \\
        r_0 &\mapsto \hat{r}(2\pi,r_0,I)
        ,
    \end{align}
    which expands to \(P(r_0) = r_0 + IS(r_0) + R(2\pi,r_0,I)\).
    In other words, we can understand the evolution of \(\hat{r}\) by studying
    the dynamics of the discrete-time system
    \begin{equation}\label{eqn:r-discrete}
        r_{n+1} = r_n + IS(r_n) + R(2\pi,r_n,I)
        ,
    \end{equation}
    with initial condition \(r_0\).
    Here, \(r_n\) corresponds to the distance along the line segment
    \(L := ]\underline{r},\overline{r}[ \, \times \{0\}\)
    when an orbit of \((r,\theta)\) hits \(L\) for the the \(n^\text{th}\) time,
    assuming the acrobot was initialized at \((r,\theta) = (r_0,0)\).

@@

Next we compute
\(b(\theta,r_0) = (\partial g/\partial I)(\theta, r_0, 0) = K a(\theta,r_0)\),
where
\begin{align*}
    K &:= \frac{\bar{q}_a \sqrt{30m^2g l^3}}{15}
    , \\
    a(\theta,r_0) &:= \frac{
        r_0 |\sin(\theta)| \left(
        5 c_{r_0} \cos(r_0 c_\theta) - 8 \cos(r_0c_\theta)^2 + 3
    \right)
    }{
    \sin(r_0)\sqrt{\cos(r_0c_\theta) - c_{r_0}}
    }
    .
\end{align*}
Notice that \(K\) is a positive constant which depends only on
\(m\),\(g\),\(l\), and \(\bar{q}_a\), and \(a(\theta,r_0)\) is \(2\pi\)-periodic
in \(\theta\).
This means that there exists \(\epsilon > 0\) such that 
\(S(r_0) := \int_0^{2\pi} b(\sigma,r_0)d\sigma \geq \epsilon\) at \(r_0\)
if and only if there exists \(\delta > 0\) such that 
\(Q(r_0) := \int_0^{2\pi} a(\sigma,r_0) d\sigma \geq \delta\).
Since \(a(\theta,r_0)\) is adimensional, we can numerically compute \(Q(r_0)\)
once for all acrobots.
The plot of \(Q(r_0)\) when \(r_0 \in [10^{-10}, \pi - 10^{-3}]\) is shown in
Figure \ref{fig:acrobot-Q}.
We see that it is strictly positive and monotonically increasing, with an
asymptote at \(r_0 = \pi\).
Simulations with smaller \(r_0\) result in an infinite integral error due to a
division by zero, yet we believe the \(Q(r_0)\) is in fact positive for all 
\(r_0 \in \, ]0,\pi[\).
\begin{figure}
    \centering
    % TODO: Show this plot
    %\includegraphics[width=0.8\linewidth]{}
    \caption{The plot of \(Q(r_0)\).}
    \label{fig:acrobot-Q}
\end{figure}

Unfortunately there is no \(\delta\) such that \(Q(r_0) \geq \delta\)
everywhere on \(]0,\pi[\), though this \(\delta\) does exist 
on \(]\omega,\pi[\) for each fixed \(\omega > 0\).
This means all conditions of Lemma \ref{lemma:energy-regulation-D}
are satisfied on the set 
\[
    D_\omega = \mathcal{O}_1 \backslash 
    \{(q_u,p_u) \in \SxR \mid E(q_u,p_u) \leq E(\omega,0)\}
    .
\]
Therefore, for each \(\omega \in \, ]0,\pi[\) there exists some 
\(I_1(\omega) > 0\) such that the acrobot gains/loses energy everywhere on
\(D_\omega\).
To prove energy gain on all of \(\mathcal{O}_1\), we will show the origin is a
repeller so that all orbits near the origin flow away from the origin into
\(D_\omega\) (for small enough \(\omega\)) and continue gaining energy
thereafter.

Recall the discrete-time pseudo-radius dynamics \eqref{eqn:r-discrete} from the
proof of Lemma \ref{lemma:energy-regulation-D}. 
To prove the origin in the \((q_u,p_u)\)-plane is a repeller, we need simply
prove the linearization of \eqref{eqn:r-discrete} is greater than \(1\) at 
\(r_0 = 0\).
Linearizing these discrete-time dynamics is equivalent to linearizing the
Poincar\'{e} map \eqref{eqn:poincare-map} to get
\[
    P^\prime(0) = 1 + I K Q^\prime(0) + R^\prime(2\pi,0,I)
    ,
\]
where prime denotes differentiation with respect to \(r\).
MATLAB symbolic computations reveal that
\begin{equation*}
    \lim \limits_{r_0 \to 0^+}
    \pdiff{a}{r}(\theta,r_0) 
    = -\frac{\sqrt{2}}{2} \left(11\sin(\theta)^2 - 6\right)
    ,
\end{equation*}
which means
\begin{align*}
    Q^\prime(0) = \int \limits_0^\pi \pdiff{a}{\mu}(\sigma,0) d\sigma
    = \frac{\pi}{2\sqrt{2}} > 0
    .
\end{align*}
Since the remainder term \(R\) is \(O(I^2)\), its partial derivative 
\(R^\prime\) is also \(O(I^2)\).
Hence, it can be written in the form
\(I^2 \tilde{R}(I)\) where \(\tilde{R}(I)\) is smooth and zero at \(I = 0\).
Thus, there exists \(I_0 > 0\) such that 
\(I_0 K Q^\prime(0) + (I_0)^2 \tilde{R}(I) > 0\), which means
\(P^\prime(0) > 1\) for all \(I \in \, ]0, I_0]\).
We conclude that the origin is a repeller, which means there exists some
(unknown) \(\mu > 0\) where the interval \(]0,\mu]\) is negatively
invariant for \eqref{eqn:r-discrete}, and all solutions starting in this
interval flow through \(r = \mu\).
This in turn implies that all orbits starting in \(\mathcal{O}_1\) near the
origin will gain energy up to \(E(\mu,0)\).
We know there exists \(I_1(\mu)\) so that the acrobot gains energy on \(D_\mu\).
Taking \(I^\star = \min\{I_0,I_1(\mu)\}\) means the acrobot will gain energy on
\(\mathcal{O}_1\) for each \(I \in \, ]0,I^\star]\).
If \(I < 0\), then \(P^\prime(0) < 1\) which implies the origin is an
asymptotically stable equilibrium.
Since the acrobot now loses energy on \(D_\mu\), all orbits in \(\mathcal{O}_1\)
lose energy and converge to the origin for \(I \in [-I^\star,0[\).
